---
title: "Kaggle Modeling"
author: "Warner Rose"
date: "May 9, 2017"
output: html_document
---
### xgboost model

Model use few macro variables, most listing-related variables, and some new features. The macro data has to be joined to the train and test data based on overlapping time stamps

```{r}
# selecting macro variables
macro_preds <- macro[c("gdp_quart","ppi",
                       "rent_price_4.room_bus",
                       "rent_price_3room_bus",
                       "rent_price_2room_bus",
                       "rent_price_1room_bus",
                       "rent_price_3room_bus",
                       "rent_price_2room_bus",
                       "rent_price_1room_bus",
                       "apt_rent_business",
                       "apt_rent_eco",
                       "micex",
                       "timestamp")]

xgb_train_df <- train %>% select(-id, -ID_metro)

xgb_train_df <- left_join(xgb_train_df, macro_preds, by="timestamp") %>%
  select(-timestamp)





xgb_train_df <- xgb_train_df %>% filter(!is.na(price_doc)) 

xgb_train_price <- xgb_train_df$price_doc

xgb_train_matrix <- as.matrix(xgb_train_df %>% select(-price_doc))


xgb_train_matrix <- as(xgb_train_matrix, "dgCMatrix")
xgb_train_matrix <- xgb.DMatrix(xgb_train_matrix, label = xgb_train_price)
```

```{r}
glimpse(xgb_train_df)
```

Try a few tuning parmeters for xgboosted trees

```{r, eval=FALSE}
#set.seed(63)

boost1 <- xgboost(data = xgb_train_matrix,
                  max_depth = 15, eta = 0.05, nrounds = 100,
                  objective = "reg:linear")

```

```{r, eval=FALSE}
boost2 <- xgboost(data = xgb_train_matrix, 
                  max_depth = 20, eta = 0.05, nrounds = 100,
                  objective = "reg:linear")
```


the train rmse looks abyssmal

```{r eval=FALSE}
boost1
boost2
```


Prepare test data and generate predictions

```{r eval=FALSE}
xgb_test_df <- test %>% select(-id, -ID_metro)
xgb_test_df <- left_join(xgb_test_df, macro_preds, by = "timestamp") %>% select(-timestamp)

xgb_test_matrix <- as.matrix(xgb_test_df)
xgb_test_matrix <- as(xgb_test_matrix, "dgCMatrix")

preds1 <- predict(boost1, xgb_test_matrix)
preds2 <- predict(boost2, xgb_test_matrix)

renderPlot(hist(preds1, main="Sanity check: histogram of 'boost1' predictions"))
renderPlot(hist(preds2, main="Sanity check: histogram of 'boost2' predictions"))
```

My gut tells me *boost2* is probably the best submission... ¯\_(ツ)_/¯

```{r eval=FALSE}
submission <- data.frame(
  id = test$id,
  price_doc = preds2
)

head(submission)
```

```{r, eval=FALSE}
write.csv(submission, "submission1.csv", row.names=F)
```
