---
title: "Kaggle EDA"
output: html_document
runtime: shiny
---

This R Markdown file contains the EDA  and model-fitting for the Kaggle competition [predicting Russian house prices](https://www.kaggle.com/c/sberbank-russian-housing-market#evaluation). Markdown source code is on [github](https://github.com/rampje/Russian-Housing-Market).

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(shiny)
library(dygraphs)
library(ggplot2)
library(xgboost)
library(dygraphs)
library(lubridate)
library(xts)
library(Matrix)
```


Feature generating function to be applied to both the training and test data

```{r}
generate_features <- function(df){
  
  df$timestamp <- as.Date(df$timestamp)
  df$day <- weekdays(df$timestamp)
  df$month <- months(df$timestamp)
  df$year_month <- df$timestamp %>% as.yearmon %>% as.Date

  # generate features specifically for train/test df
  if("gdp_quart" %in% names(df)){
  
    # aggregated rent price features
    df$apt_rent_business <- macro[grep("room_bus",names(macro))] %>% rowMeans(na.rm = TRUE)
    df$apt_rent_eco <- macro[grep("room_eco",names(macro))] %>% rowMeans(na.rm = TRUE)
    
  } else{
    
    df$volume <- rep(1, nrow(df))
    df$month_vol <- with(df, ave(volume, year_month, FUN = function(x) sum(x)))
    df$rel_floor <- df$floor / df$max_floor
    df$rel_living <- df$life_sq / df$full_sq
    df$rel_kitch <- df$kitch_sq / df$full_sq
    
    # convert categorical variables to dummies
    df$culture_objects_top_25 <- ifelse(df$culture_objects_top_25 == "yes", 1,0)
    df$thermal_power_plant_raion <- ifelse(df$thermal_power_plant_raion == "yes", 1,0)
    df$oil_chemistry_raion <- ifelse(df$culture_objects_top_25 == "yes", 1,0)
    df$radiation_raion <- ifelse(df$culture_objects_top_25 == "yes", 1,0)
    df$railroad_terminal_raion <- ifelse(df$railroad_terminal_raion == "yes",1,0)
    df$big_market_raion <- ifelse(df$big_market_raion == "yes",1,0)
    df$nuclear_reactor_raion <- ifelse(df$nuclear_reactor_raion == "yes",1,0)
    df$detention_facility_raion <- ifelse(df$detention_facility_raion == "yes",1,0)
    df$incineration_raion <- ifelse(df$incineration_raion == "yes", 1, 0)
    df$big_road1_1line <- ifelse(df$big_road1_1line == "yes", 1,0)
    df$railroad_1line <- ifelse(df$railroad_1line == "yes", 1,0)
    

  }
  
  df
}
```

Read in data

```{r}
train <- read.csv("train.csv", stringsAsFactors = FALSE)
macro <- read.csv("macro.csv", stringsAsFactors = FALSE)
test <- read.csv("test.csv", stringsAsFactors = FALSE)
```

Transform columns and generate new features

```{r, message=FALSE}
train <- generate_features(train)
macro <- generate_features(macro)
test <- generate_features(test)
```

### Analysis of time series

```{r, message=FALSE, warning=FALSE, include=FALSE}
price <- train[c("timestamp", "price_doc")]

series_shell <- seq(min(price$timestamp),
                    max(price$timestamp), by = 1)
series_shell <- data.frame("timestamp"=series_shell)

price <- price %>%
          group_by(timestamp) %>%
          summarise_each(funs(mean(., na.rm=TRUE)))

price$timestamp <- as.Date(price$timestamp)

price_qnt <- quantile(price$price_doc, probs = 0.98)
  
price <- price %>% filter(price_doc < price_qnt)
price <- full_join(series_shell, price)

price_ts <- xts(price[,-1], 
                order.by = price[,1])

```


```{r, message=FALSE, warning=FALSE, include=FALSE}
volume <- train["timestamp"]
volume$volume <- rep(1, nrow(volume))

volume$timestamp <- as.Date(volume$timestamp)

volume <- volume %>%
            group_by(timestamp) %>%
            summarise_each(funs(sum))

volume <- full_join(series_shell, volume, by="timestamp")
volume$timestamp <- as.Date(volume$timestamp,
                            format="%Y-%m-%d")

volume_ts <- xts(volume[,-1],
                 order.by = volume[,1])
```


```{r, echo=FALSE, message=FALSE, warning=FALSE}
selectInput("macro", "select macro variable:", 
          names(macro)[names(macro) != "timestamp"])

selectInput("diff", "series difference lag:",
            c(0,1,2), selected = 0)

#selectInput("transform", "select transformation:",
#            c("none","log","square root", "first #difference"), selected = "none")

renderDygraph({
selection_col <- names(macro)[names(macro) == input$macro]
plot_array <- macro[c("timestamp", selection_col)]

plot_array <- xts(plot_array[,-1],
                  order.by = as.Date(plot_array[,1], "%Y-%m-%d"))

if(input$diff %in% 1:2){
  plot_array <- diff(plot_array, lag = input$diff)
}

plot_array <- cbind(price_ts, plot_array)
dygraph(plot_array, main=paste("price_doc and",input$macro)) %>%
  dyRangeSelector(height = 40,
                      dateWindow = c("2011-07-01", "2015-9-01")) %>%
  dySeries("..2", axis = "y2") %>%
  dyRoller()
})

renderDygraph({
  dygraph(cbind(price_ts, volume_ts), main="price_doc and volume") %>% dyRangeSelector() %>%
    dySeries("..2", axis="y2") %>%
    dyOptions(fillGraph=TRUE, strokeWidth = 1)
})

```



### Histograms of numeric variables

```{r, echo = FALSE}
train_classes <- map_chr(train, class)
train_numerics <- train_classes[train_classes %in% c("integer","numeric")] %>% names

sliderInput("quantile", "select quantiles:", 
            min = 0, max = 100, value = c(0, 98))

renderPlot({
  qnt <- as.numeric(quantile(train$price_doc, probs = input$quantile/100))
  
  train$price_doc[train$price_doc <= qnt] %>% 
    hist(main="Histogram of price_doc")
})

selectInput("train_var", "select variable to compare with `price_doc`:", 
          train_numerics[train_numerics != "id"])

renderPlot({
  index <- which(names(train) == input$train_var)
  plot_data <- as.numeric(unlist(train[index]))
  
  qnt <- as.numeric(quantile(plot_data, probs = input$quantile/100,
                             na.rm = TRUE))
  
  plot_data[plot_data <= qnt] %>% 
    hist(main=paste("Histogram of",input$train_var))
})

renderPlot({
  y_qnt <- as.numeric(quantile(train$price_doc, probs = input$quantile/100))
  
  y <- train$price_doc[train$price_doc <= y_qnt]

  index <- which(names(train) == input$train_var)
  plot_data <- as.numeric(unlist(train[index]))
  
  x <- plot_data[train$price_doc <= y_qnt]
  
  plot(x,y)
})
```

### xgboost model

Model use few macro variables, most listing-related variables, and some new features. The macro data has to be joined to the train and test data based on overlapping time stamps

```{r}
# selecting macro variables
macro_preds <- macro[c("gdp_quart","ppi",
                       "rent_price_4.room_bus",
                       "rent_price_3room_bus",
                       "rent_price_2room_bus",
                       "rent_price_1room_bus",
                       "rent_price_3room_bus",
                       "rent_price_2room_bus",
                       "rent_price_1room_bus",
                       "apt_rent_business",
                       "apt_rent_eco",
                       "micex",
                       "timestamp")]

xgb_train_df <- train %>% select(-id, -ID_metro)

xgb_train_df <- left_join(xgb_train_df, macro_preds, by="timestamp") %>%
                  select(-timestamp)





xgb_train_df <- xgb_train_df %>% filter(!is.na(price_doc)) 

xgb_train_price <- xgb_train_df$price_doc

xgb_train_matrix <- as.matrix(xgb_train_df %>% select(-price_doc))


xgb_train_matrix <- as(xgb_train_matrix, "dgCMatrix")
xgb_train_matrix <- xgb.DMatrix(xgb_train_matrix, label = xgb_train_price)
```

```{r}
glimpse(xgb_train_df)
```

Try a few tuning parmeters for xgboosted trees

```{r, eval=FALSE}
#set.seed(63)

boost1 <- xgboost(data = xgb_train_matrix,
                 max_depth = 15, eta = 0.05, nrounds = 100,
                 objective = "reg:linear")

```

```{r, eval=FALSE}
boost2 <- xgboost(data = xgb_train_matrix, 
                  max_depth = 20, eta = 0.05, nrounds = 100,
                  objective = "reg:linear")
```


the train rmse looks abyssmal

```{r eval=FALSE}
boost1
boost2
```


Prepare test data and generate predictions

```{r eval=FALSE}
xgb_test_df <- test %>% select(-id, -ID_metro)
xgb_test_df <- left_join(xgb_test_df, macro_preds, by = "timestamp") %>% select(-timestamp)

xgb_test_matrix <- as.matrix(xgb_test_df)
xgb_test_matrix <- as(xgb_test_matrix, "dgCMatrix")

preds1 <- predict(boost1, xgb_test_matrix)
preds2 <- predict(boost2, xgb_test_matrix)

renderPlot(hist(preds1, main="Sanity check: histogram of 'boost1' predictions"))
renderPlot(hist(preds2, main="Sanity check: histogram of 'boost2' predictions"))
```

My gut tells me *boost2* is probably the best submission... ¯\_(ツ)_/¯

```{r eval=FALSE}
submission <- data.frame(
  id = test$id,
  price_doc = preds2
)

head(submission)
```

```{r, eval=FALSE}
write.csv(submission, "submission1.csv", row.names=F)
```

